---
title       : An app to predict timber volume
subtitle    : 
author      : Rachel Karpman
job         : 
framework   : io2012   # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [bootstrap]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

<style>
.title-slide {
  background-color: #FAF0E6;
  color: #654321;
}

.title-slide hgroup > h1 {
    color: #006400;
}

slide:not(.segue) {
    background-color: #FAF0E6;
    color: #654321;
}


slide:not(.segue) h2{
    color: #006400;
}

</style>

## Introduction

* Estimating the volume of trees is important for anyone who harvests timber, either for their own use or commercially.

    + Since trees have complex shapes, measuring their volume directly is difficult

* `Timber Volume` is a `Shiny` app that predicts a tree's volume from its height and girth.

### The `trees` data set

* We build a model using the `trees` data set, which contains the height, girth, and timber volume of 31 felled cherry trees.

    + Girth is diameter of the tree trunk, measured at 4 feet 6 inches above the ground
    + Height is measured in feet, girth is in inches, and volume is in cubic feet.

---
## Modeling timber volume

* We model timber volume as a linear function of `Height * Girth^2`.
    
    + This is similar to the volume formula for cones and cylinders.

```{r}
model <- lm(Volume ~ I(Girth^2 * Height), data = trees)
```

* For example, if girth is 10 inches and height is 70 feet, we can use this model to predict timber volume using the code below.

```{r}
newVolume <- predict(model, data.frame(Girth = 10, Height = 70))
newVolume
```

We get an estimated volume of `r round(newVolume, 2)` cubic feet.

--- 

## Using the app

* The `Timber Volume` app has a simple interface.

    + A `Documentation` tab contains instructions for new users.

* The user selects the desired girth and height with sliders.

    + Predicted volume is calculated and displayed automatically

* The app also displays a plot of height versus girth for each point in the `trees` dataset, and for user's tree.

    + Point size represents volume (observed or predicted).

--- 

## A data visualization

* The figure shows similar to the one displayed by the Timber Volume app.

    + The brown circle represents a predicted tree with height 70 feet and girth 10 inches, and predicted volume `r round(newVolume, 2)` cubic feet.

```{r echo = FALSE, warnings = FALSE, message = FALSE, fig.height = 5, fig.align = "center"}
library(ggplot2)
library(plotly)
newPoint <- data.frame(Height = 70, Girth = 10, Volume = newVolume)
p <- ggplot(trees, aes(x = Height, 
                       y = Girth, 
                       size = Volume, 
                       text = paste('Height: ', Height,
                                    '<br> Girth: ', Girth,
                                    '<br>Volume: ', Volume))) +
    geom_point(color = "green4", alpha = 0.4) +
    geom_point(data = newPoint, color = "sienna4", alpha = 0.8) +
    xlim(60, 90) +
    ylim(5, 25)
ggplotly(p, tooltip = "text")
```




